### 2021-02-21 paper.tex
Corrections. Submission version.

Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own~\cite{ontologybased}.
There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges.
but they do not fit our requirements~\cite{visualizationoflargeontologies}.
Our contribution is SNIK Graph, a web application\footnote{\url{http://www.snik.eu/graph}} that offers Linked Data visualization for both classes and instances and according to the requirements of SNIK and its users.
%The graph is visualized using the Cytoscape.js~\cite{cytoscapejs} library with the force-directed Euler layout. %, see%~\cref{fig:snik-graph-overview}.
With the thousands of classes of SNIK, graph-based visualization suffers from overplotting.
Thus, SNIK Graph offers several options to select and layout subgraphs, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
Users can also iteratively explore SNIK starting at a single class using neighbourhood and path operations. %(see \cref{fig:snik-graph-circle-star})
% and subsequently the neighbourhood of selected neighbours of the previous step.
%SNIK Graph can also calculate the shortest path to between classes.
%Path and neighourhood operations are joined in the \enquote{spider worm}
%(see \cref{fig:snik-graph-spiderworm})
%, which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.

@@

\subsubsection*{Graph Rendering}
Due to the low performance of rendering on an HTML canvas, the frame rate can drop significantly below a fluent 60 frames per second (FPS).
While Cytoscape.js buffers already rendered parts, this mechanism suffers when moving around at medium zoom levels.
The frame rate is especially low on large resolutions such as 4k (3840x2160) and on browsers other than Google Chrome / Chromium, where the minimum frame rate drops to as low as 8 FPS on an Intel i9-10900k at 4k on Mozilla Firefox 86 and 36 FPS on Chromium 88 on Arch Linux.
We attribute the large difference between Chrome / Chromium and Firefox to the different JavaScript engines.
While SNIK Graph does not require perfectly smooth motion and wild movements are not a common usage pattern, stuttering is still frustrating to users especially on less performant CPUs and browsers other than Chrome and contrary to our goals of minimizing friction for users. 
Due to the enormous power of modern GPUs, implementing an OpenGL-based renderer for Cytoscape.js has the potential to dramatically increase render speed.
High framerates would also allow the study of nausea-free virtual reality presentation for Linked Data visualization.

>>>>

Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own~\cite{ontologybased}.
There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges, but they do not fit our requirements~\cite{visualizationoflargeontologies}.
Our contribution is SNIK Graph, a web application\footnote{\url{http://www.snik.eu/graph}} that offers Linked Data visualization for both classes and instances and according to the requirements of SNIK and its users.
%The graph is visualized using the Cytoscape.js~\cite{cytoscapejs} library with the force-directed Euler layout. %, see%~\cref{fig:snik-graph-overview}.
With the thousands of classes of SNIK, graph-based visualization suffers from overplotting.
Thus, SNIK Graph offers several options to select and layout subgraphs, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
Users can also iteratively explore SNIK starting at a single class using neighbourhood and path operations. %(see \cref{fig:snik-graph-circle-star})
% and subsequently the neighbourhood of selected neighbours of the previous step.
%SNIK Graph can also calculate the shortest path to between classes.
%Path and neighourhood operations are joined in the \enquote{spider worm}
%(see \cref{fig:snik-graph-spiderworm})
%, which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.

@@

\subsubsection*{Graph Rendering}
Due to the low performance of rendering on an HTML canvas, the frame rate can drop significantly below a fluent 60 frames per second (FPS).
While Cytoscape.js buffers already rendered parts, this mechanism suffers when moving around at medium zoom levels.
The frame rate is especially low on large resolutions such as 4k (3840x2160) and on browsers other than Google Chrome / Chromium, where the minimum frame rate drops to as low as 8 FPS on an Intel i9-10900k at 4k on Mozilla Firefox 86 and 36 FPS on Chromium 88 on Arch Linux.
We attribute the large difference between Chrome / Chromium and Firefox to the different JavaScript engines.
While SNIK Graph does not require perfectly smooth motion and wild movements are not a common usage pattern, stuttering is still frustrating to users especially on less performant CPUs and browsers other than Chrome and contrary to our goal of minimizing friction for users. 
Due to the enormous power of modern GPUs, implementing an OpenGL-based renderer for Cytoscape.js has the potential to dramatically increase render speed.
High framerates would also allow the study of nausea-free virtual reality presentation for Linked Data visualization.
